Dominance Regions for Imaginary $\bfg$-vectors

### 2021-06-25 dominance_regions.tex
Minor edits

\newcommand{\Aut}{\operatorname{Aut}}
\newcommand{\diag}{\operatorname{diag}}
\newcommand{\Id}{\operatorname{Id}}
\newcommand{\into}{\hookrightarrow}
\newcommand{\obeta}{{\overline{\beta}}}
\newcommand{\oi}{{\overline{\imath}}}
\newcommand{\ot}{{\overline{t}}}

@@

  \section{Lattice Mutations}
  Let $B=(b_{ij})$ be an $n\times n$ skew-symmetrizable integer matrix, i.e. there exists a diagonal integer matrix $D=\diag(d_1,\ldots,d_n)$ so that $DB$ is skew-symmetric, more precisely we have $d_i b_{ij}=-d_j b_{ji}$.
  For $b\in\ZZ$, write $[b]_+=max(b,0)$.
  Given a sign $\varepsilon\in\{\pm\}$ and $k\in\{1,\ldots,n\}$, define a matrix $E_{k,\varepsilon}=(e_{ij})$ with
  \begin{equation}
    \label{eq:left mutation matrix}
    e_{ij}=\begin{cases} 1 & \text{if $i=j\ne k$;}\\ -1 & \text{if $i=j=k$;}\\ [\varepsilon b_{ik}]_+ & \text{if $i\ne j=k$;}\\ 0 & \text{otherwise;} \end{cases}
  \end{equation}
  and a matrix $F_{k,\varepsilon}=(f_{ij})$ with
  \begin{equation}
    \label{eq:right mutation matrix}
    f_{ij}=\begin{cases} 1 & \text{if $k\ne i=j$;}\\ -1 & \text{if $k=i=j$;}\\ [-\varepsilon b_{kj}]_+ & \text{if $k=i\ne j$;}\\ 0 & \text{otherwise.} \end{cases}
  \end{equation}
  The index $k\in\{1,\ldots,n\}$ also determines a new skew-symmetrizable matrix $\mu_kB=(b'_{ij})$ given by
  \[ b'_{ij}=\begin{cases} -b_{ij} & \text{if $i=k$ or $j=k$;}\\ b_{ij}+[b_{ik}]_+b_{kj}+b_{ik}[-b_{kj}]_+ & \text{otherwise.} \end{cases} \]
  One easily observes that $\mu_k B$ is again skew-symmetrizable using the same matrix $D$.
  Note that $\mu_k B=E_{k,\varepsilon} B F_{k,\varepsilon}$ for $\varepsilon=\pm 1$, the case $\varepsilon=1$ being obvious from the definitions and the case $\varepsilon=-1$ following from the identity $b_{ij}=[b_{ij}]_+-[-b_{ij}]_+$.

@@

  (DR: I changed to $\RR$ because it didn't seem like we really cared about integer points at the moment, when we try to link up with minors this will need to be added.)
  For $t\in\TT_n$, define $\phi^t_k:\RR^n\to\RR^n$ as the piecewise-linear map
  \[\phi^t_k(\lambda)=\begin{cases} E^t_{k,+}\lambda & \text{if $\lambda_k\ge0$;}\\ E^t_{k,-}\lambda & \text{if $\lambda_k<0$;} \end{cases}\]
  where the entries of $E^t_{k,\varepsilon}$ is given by \eqref{eq:left mutation matrix} with $b^t_{ij}$ in place of $b_{ij}$.
  We leave it as an exercise for the reader to check that $(\phi^t_k)^{-1}=\phi^{t'}_k$ whenever $t,t'\in\TT_n$ are joined by an edge labeled $k$.
  For $t\in\TT_n$, define the piecewise-linear automorphisms $\phi_t:\RR^n\to\RR^n$ by
  \begin{itemize}
    \item $\phi_{t_0}=\Id$;
    \item if $t,t'\in\TT_n$ are joined by an edge labeled $k$, then $\phi_{t'}=\phi^t_k \phi_t$.
  \end{itemize}
  For $\lambda\in\RR^n$, define the \emph{$B_t$-cone} $\cC_t(\lambda):=\lambda+B_t\cdot\RR_{\ge0}^n$ pointed at $\lambda$.
  \begin{definition}
    For $\lambda\in\RR^n$, define the \emph{dominance region} 
    \[ \cP(\lambda) = \bigcap_{t\in\TT_n} \phi_t^{-1} \cC_t(\phi_t \lambda). \]
    When $\mu\in\cP(\lambda)$, we say \emph{$\lambda$ dominates $\mu$}.
  \end{definition}
  Our goal is to understand the dominance polytopes $\cP(\lambda)$ for all $\lambda$.
  \begin{lemma}
    The sets $\cP(\lambda)$ are polytopes.
  \end{lemma}

@@

  The dominance polygon pointed at $\lambda=(\lambda_1,\lambda_2)$ is the region defined by the following six inequalities:
  \begin{align}
    -\lambda_1-\frac{b c+\sqrt{b c (b c-4)}}{2 c}\lambda_2+x+\frac{b c+\sqrt{b c (b c-4)}}{2 c}y\geq0
    \\
    \frac{b c+\sqrt{b c (b c-4)}}{2 b}\lambda_1+\lambda_2-\frac{b c+\sqrt{b c (b c-4)}}{2 b}x-y\geq0
    \\
    -\frac{b c+\sqrt{b c (b c-4)}}{2 b}\lambda_1 - \frac{b c+\sqrt{b c (b c-4)}-2}{2}\lambda_2 + \frac{b c+\sqrt{b c (b c-4)}}{2 b}x-y\geq0
    \\
    \frac{b c+\sqrt{b c (b c-4)}-2}{2}\lambda_1+\frac{b c+\sqrt{b c (b c-4)}}{2 c}\lambda_2+x-\frac{b c+\sqrt{b c (b c-4)}}{2 c}y\geq0
    \\
    -\frac{b c+\sqrt{b c (b c-4)}}{2 b}\lambda_1-\frac{b c+\sqrt{b c (b c-4)}-2}{2}\lambda_2-\frac{b c-\sqrt{b c (b c-4)}}{2 b}x-y\geq0
    \\
    \frac{b c+\sqrt{b c (b c-4)}-2}{2}\lambda_1+\frac{b c+\sqrt{b c (b c-4)}}{2 c}\lambda_2+x+\frac{b c-\sqrt{b c (b c-4)}}{2 c}y\geq0
  \end{align}
  Are any of these better forms?
  \begin{align}
    (x-\lambda_1)+\frac{b c+\sqrt{b c (b c-4)}}{2 c}(y-\lambda_2)\geq0
    \\
    \nonumber \iff \frac{b c-\sqrt{b c (b c-4)}}{2 b}(x-\lambda_1)+(y-\lambda_2)\geq0
    \\
    -\frac{b c+\sqrt{b c (b c-4)}}{2 b}(x-\lambda_1)-(y-\lambda_2)\geq0
    \\
    \nonumber \iff -(x-\lambda_1)-\frac{b c-\sqrt{b c (b c-4)}}{2 c}(y-\lambda_2)\geq0
    \\
    -\frac{b c+\sqrt{b c (b c-4)}}{2}\lambda_2 + \frac{b c+\sqrt{b c (b c-4)}}{2 b}(x-\lambda_1)-(y-\lambda_2)\geq0
    \\
    \nonumber \iff -b\lambda_2 + (x-\lambda_1)-\frac{b c-\sqrt{b c (b c-4)}}{2 c}(y-\lambda_2)\geq0
    \\
    \frac{b c+\sqrt{b c (b c-4)}}{2}\lambda_1+(x-\lambda_1)-\frac{b c+\sqrt{b c (b c-4)}}{2 c}(y-\lambda_2)\geq0
    \\
    \nonumber \iff c\lambda_1+\frac{b c-\sqrt{b c (b c-4)}}{2 b}(x-\lambda_1)-(y-\lambda_2)\geq0
    \\
    -c\lambda_1-\frac{b c+\sqrt{b c (b c-4)}}{2}\lambda_2-\frac{b c-\sqrt{b c (b c-4)}}{2 b}(x-\lambda_1)-(y-\lambda_2)\geq0
    \\
    \frac{b c+\sqrt{b c (b c-4)}}{2}\lambda_1+b\lambda_2+(x-\lambda_1)+\frac{b c-\sqrt{b c (b c-4)}}{2 c}(y-\lambda_2)\geq0
  \end{align}
  Another proposal (I also changed the order of the equations)
  \begin{align}
    \frac{b c-\sqrt{b c (b c-4)}}{2 b}(x-\lambda_1)+(y-\lambda_2) & \geq0
    \\
    \frac{b c+\sqrt{b c (b c-4)}}{2 b}(x-\lambda_1)-(y-\lambda_2) & \geq \frac{b c+\sqrt{b c (b c-4)}}{2}\lambda_2 
    \\
    -\frac{b c-\sqrt{b c (b c-4)}}{2 b}(x-\lambda_1)-(y-\lambda_2)& \geq c\lambda_1+\frac{b c+\sqrt{b c (b c-4)}}{2}\lambda_2
    \\
    -(x-\lambda_1)-\frac{b c-\sqrt{b c (b c-4)}}{2 c}(y-\lambda_2)& \geq0
    \\
    (x-\lambda_1)-\frac{b c+\sqrt{b c (b c-4)}}{2 c}(y-\lambda_2) & \geq -\frac{b c+\sqrt{b c (b c-4)}}{2}\lambda_1
    \\
    (x-\lambda_1)+\frac{b c-\sqrt{b c (b c-4)}}{2 c}(y-\lambda_2) & \geq -\frac{b c+\sqrt{b c (b c-4)}}{2}\lambda_1-b\lambda_2
  \end{align}

>>>>

\newcommand{\Aut}{\operatorname{Aut}}
\newcommand{\diag}{\operatorname{diag}}
\newcommand{\Id}{\operatorname{Id}}
\renewcommand{\Im}{\operatorname{Im}}
\newcommand{\into}{\hookrightarrow}
\newcommand{\obeta}{{\overline{\beta}}}
\newcommand{\oi}{{\overline{\imath}}}
\newcommand{\ot}{{\overline{t}}}

@@

  \section{Lattice Mutations}
  Let $B=(b_{ij})$ be an $n\times n$ skew-symmetrizable integer matrix, i.e. there exists a diagonal integer matrix $D=\diag(d_1,\ldots,d_n)$ so that $DB$ is skew-symmetric, more precisely we have $d_i b_{ij}=-d_j b_{ji}$.
  For $b\in\ZZ$, write $[b]_+=max(b,0)$.
  Given a sign $\varepsilon\in\{\pm\}$ and $k\in\{1,\ldots,n\}$, define a matrix $E_{k,\varepsilon}=(e_{ij})$ with
  \begin{equation}
    \label{eq:left mutation matrix}
    e_{ij}=\begin{cases} 1 & \text{if $i=j\ne k$;}\\ -1 & \text{if $i=j=k$;}\\ [\varepsilon b_{ik}]_+ & \text{if $i\ne j=k$;}\\ 0 & \text{otherwise;} \end{cases}
  \end{equation}
  and a matrix $F_{k,\varepsilon}=(f_{ij})$ with
  \begin{equation}
    \label{eq:right mutation matrix}
    f_{ij}=\begin{cases} 1 & \text{if $k\ne i=j$;}\\ -1 & \text{if $k=i=j$;}\\ [-\varepsilon b_{kj}]_+ & \text{if $k=i\ne j$;}\\ 0 & \text{otherwise.} \end{cases}
  \end{equation}
  Observe that $E^2_{k,\varepsilon}=\Id$ and $F^2_{k,\varepsilon}=\Id$ for any $k$ and any choice of $\varepsilon$.

  The index $k\in\{1,\ldots,n\}$ also determines a new skew-symmetrizable matrix $\mu_kB=(b'_{ij})$ given by
  \[ b'_{ij}=\begin{cases} -b_{ij} & \text{if $i=k$ or $j=k$;}\\ b_{ij}+[b_{ik}]_+b_{kj}+b_{ik}[-b_{kj}]_+ & \text{otherwise.} \end{cases} \]
  One easily observes that $\mu_k B$ is again skew-symmetrizable using the same matrix $D$.
  \begin{remark}
    Note that $\mu_k B=E_{k,\varepsilon} B F_{k,\varepsilon}$ for $\varepsilon=\pm 1$, the case $\varepsilon=1$ being obvious from the definitions and the case $\varepsilon=-1$ following from the identity $b_{ij}=[b_{ij}]_+-[-b_{ij}]_+$.
  \end{remark}

@@

  (DR: I changed to $\RR$ because it didn't seem like we really cared about integer points at the moment, when we try to link up with minors this will need to be added.)
  For $t\in\TT_n$, define $\phi^t_k:\RR^n\to\RR^n$ as the piecewise-linear map
  \[\phi^t_k(\lambda)=\begin{cases} E^t_{k,+}\lambda & \text{if $\lambda_k\ge0$;}\\ E^t_{k,-}\lambda & \text{if $\lambda_k<0$;} \end{cases}\]
  where the entries of $E^t_{k,\varepsilon}$ are given by \eqref{eq:left mutation matrix} with $b^t_{ij}$ in place of $b_{ij}$.
  We leave it as an exercise for the reader to check that $(\phi^t_k)^{-1}=\phi^{t'}_k$ whenever $t,t'\in\TT_n$ are joined by an edge labeled $k$.
  \begin{lemma}
    Suppose $t,t'\in\TT_n$ are joined by an edge labeled $k$.
    If $\lambda\in\Im(B^t)$, then $\phi^t_k(\lambda)\in\Im(B^{t'})$.
  \end{lemma}
  \begin{proof}
    Suppose $\lambda=B^t\alpha$ for some $\alpha\in\RR^n$.
    Then 
    \[\phi^t_k(\lambda)=E^t_{k,\pm}\lambda=E^t_{k,\pm} B^t\alpha=E^t_{k,\pm} B^t F^t_{k,\pm} F^t_{k,\pm}\alpha=B^{t'} F^t_{k,\pm}\alpha\]
    so that $\phi^t_k(\lambda)\in\Im(B^{t'})$.
  \end{proof}

  For $t\in\TT_n$, define the piecewise-linear automorphisms $\phi_t:\RR^n\to\RR^n$ by
  \begin{itemize}
    \item $\phi_{t_0}=\Id$;
    \item if $t,t'\in\TT_n$ are joined by an edge labeled $k$, then $\phi_{t'}=\phi^t_k \phi_t$.
  \end{itemize}
  For $\lambda\in\RR^n$, define the \emph{$B_t$-cone} $\cC_t(\lambda):=\lambda+B_t\cdot\RR_{\ge0}^n$ pointed at $\lambda$.
  \begin{definition}
    For $\lambda\in\RR^n$, define the \emph{dominance region} 
    \[ \cP(\lambda) = \bigcap_{t\in\TT_n} \phi_t^{-1} \cC_t(\phi_t \lambda). \]
    When $\mu\in\cP(\lambda)$, we say \emph{$\lambda$ dominates $\mu$}.
  \end{definition}
  Our goal is to understand the dominance polytopes $\cP(\lambda)$ for all $\lambda$.
  \begin{lemma}
    The sets $\cP(\lambda)$ are polytopes.
  \end{lemma}

@@

  The dominance polygon pointed at $\lambda=(\lambda_1,\lambda_2)$ is the region defined by the following six inequalities:
  \begin{align}
    -\lambda_1-\frac{b c+\sqrt{b c (b c-4)}}{2 c}\lambda_2+x+\frac{b c+\sqrt{b c (b c-4)}}{2 c}y\geq0
    \\
    \frac{b c+\sqrt{b c (b c-4)}}{2 b}\lambda_1+\lambda_2-\frac{b c+\sqrt{b c (b c-4)}}{2 b}x-y\geq0
    \\
    -\frac{b c+\sqrt{b c (b c-4)}}{2 b}\lambda_1 - \frac{b c+\sqrt{b c (b c-4)}-2}{2}\lambda_2 + \frac{b c+\sqrt{b c (b c-4)}}{2 b}x-y\geq0
    \\
    \frac{b c+\sqrt{b c (b c-4)}-2}{2}\lambda_1+\frac{b c+\sqrt{b c (b c-4)}}{2 c}\lambda_2+x-\frac{b c+\sqrt{b c (b c-4)}}{2 c}y\geq0
    \\
    -\frac{b c+\sqrt{b c (b c-4)}}{2 b}\lambda_1-\frac{b c+\sqrt{b c (b c-4)}-2}{2}\lambda_2-\frac{b c-\sqrt{b c (b c-4)}}{2 b}x-y\geq0
    \\
    \frac{b c+\sqrt{b c (b c-4)}-2}{2}\lambda_1+\frac{b c+\sqrt{b c (b c-4)}}{2 c}\lambda_2+x+\frac{b c-\sqrt{b c (b c-4)}}{2 c}y\geq0
  \end{align}
  Are any of these better forms?
  \begin{align}
    (x-\lambda_1)+\frac{b c+\sqrt{b c (b c-4)}}{2 c}(y-\lambda_2)\geq0
    \\
    \nonumber \iff \frac{b c-\sqrt{b c (b c-4)}}{2 b}(x-\lambda_1)+(y-\lambda_2)\geq0
    \\
    -\frac{b c+\sqrt{b c (b c-4)}}{2 b}(x-\lambda_1)-(y-\lambda_2)\geq0
    \\
    \nonumber \iff -(x-\lambda_1)-\frac{b c-\sqrt{b c (b c-4)}}{2 c}(y-\lambda_2)\geq0
    \\
    -\frac{b c+\sqrt{b c (b c-4)}}{2}\lambda_2 + \frac{b c+\sqrt{b c (b c-4)}}{2 b}(x-\lambda_1)-(y-\lambda_2)\geq0
    \\
    \nonumber \iff -b\lambda_2 + (x-\lambda_1)-\frac{b c-\sqrt{b c (b c-4)}}{2 c}(y-\lambda_2)\geq0
    \\
    \frac{b c+\sqrt{b c (b c-4)}}{2}\lambda_1+(x-\lambda_1)-\frac{b c+\sqrt{b c (b c-4)}}{2 c}(y-\lambda_2)\geq0
    \\
    \nonumber \iff c\lambda_1+\frac{b c-\sqrt{b c (b c-4)}}{2 b}(x-\lambda_1)-(y-\lambda_2)\geq0
    \\
    -c\lambda_1-\frac{b c+\sqrt{b c (b c-4)}}{2}\lambda_2-\frac{b c-\sqrt{b c (b c-4)}}{2 b}(x-\lambda_1)-(y-\lambda_2)\geq0
    \\
    \frac{b c+\sqrt{b c (b c-4)}}{2}\lambda_1+b\lambda_2+(x-\lambda_1)+\frac{b c-\sqrt{b c (b c-4)}}{2 c}(y-\lambda_2)\geq0
  \end{align}
  Another proposal (I also changed the order of the equations)
  \begin{align}
    \frac{b c-\sqrt{b c (b c-4)}}{2 b}(x-\lambda_1)+(y-\lambda_2) & \geq0
    \\
    \frac{b c+\sqrt{b c (b c-4)}}{2 b}(x-\lambda_1)-(y-\lambda_2) & \geq \frac{b c+\sqrt{b c (b c-4)}}{2b}b\lambda_2 
    \\
    -\frac{b c-\sqrt{b c (b c-4)}}{2 b}(x-\lambda_1)-(y-\lambda_2)& \geq c\lambda_1+\frac{b c+\sqrt{b c (b c-4)}}{2b}b\lambda_2
    \\
    -(x-\lambda_1)-\frac{b c-\sqrt{b c (b c-4)}}{2 c}(y-\lambda_2)& \geq0
    \\
    (x-\lambda_1)-\frac{b c+\sqrt{b c (b c-4)}}{2 c}(y-\lambda_2) & \geq -\frac{b c+\sqrt{b c (b c-4)}}{2c}c\lambda_1
    \\
    (x-\lambda_1)+\frac{b c-\sqrt{b c (b c-4)}}{2 c}(y-\lambda_2) & \geq -\frac{b c+\sqrt{b c (b c-4)}}{2c}c\lambda_1-b\lambda_2
  \end{align}
